\thetitle

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
improved simplified 'analytical' form of error bound

Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic error bound, we first take some time to better understand the bound on $\bbk{\epsilon_{i\ell}^2}$ analytically.
To this end, we first note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
Moreover, we can also bound
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le \sum_\mu t_{\ell\mu}^2 = 1,
  &&
  \t{so}
  &&
  \bbk{\epsilon_{i\ell}^2} \le \f1n \times \min\Set{\f{2\ell+1}{d}, 1}.
\end{align}
If $d$ is even, then we can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu \le t_{\ell\mu_{\t{max}}}^2,
\end{align}
where $\mu_{\t{max}}$ is any index maximizing $t_{\ell\mu}$.
The symmetries of the Clebsch-Gordan coefficients imply that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$, so if $\mu_{\t{max}} \ne0$ (which is always true when $d$ is even because $s,\mu\notin\ZZ$), then
\begin{align}
  g_\ell\p{p} \le t_{\ell\mu_{\t{max}}}^2
  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
  \le \f12 \sum_\mu t_{\ell\mu}^2
  = \f12,
  &&
  \t{so}
  &&
  \bbk{\epsilon_{i\ell}^2}
  \le \f1n \times \min\Set{\f{2\ell+1}{d}, \f12}.
\end{align}

>>>>

Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic error bound, we first take some time to better understand the bound on $\bbk{\epsilon_{i\ell}^2}$ analytically.
To this end, we first note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
Moreover, we can also bound
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le \sum_\mu t_{\ell\mu}^2 = 1,
  &&
  \t{so}
  &&
  \bbk{\epsilon_{i\ell}^2} \le \f1n \times \min\Set{\f{2\ell+1}{d}, 1}.
\end{align}
We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering that if $\mu_{\t{max}}\ne0$, then
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le t_{\ell\mu_{\t{max}}}^2
  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
  \stackrel{\mu_{\t{max}}\ne0}{\le} \f12 \sum_\mu t_{\ell\mu}^2
  = \f12,
\end{align}
where we used the fact that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
If $\mu_{\t{max}}=0$, then
\begin{align}
  t_{\ell\mu_{\t{max}}}^2 + 2 t_{\ell\mu_{\t{min}}}^2
  = t_{\ell\mu_{\t{max}}}^2 + t_{\ell\mu_{\t{min}}}^2
  + t_{\ell,-\mu_{\t{min}}}^2
  \stackrel{\mu_{\t{max}}=0}{\le}
  \sum_\mu t_{\ell\mu}^2 = 1,
  &&
  \t{so}
  &&
  \abs{t_{\ell\mu_{\t{min}}}}
  \stackrel{\mu_{\t{max}}=0}{\le}
  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}.
\end{align}
This inequality lets us bound
\begin{align}
  \sqrt{\f{2\ell+1}{d}}\, \gamma_\ell
  = \f12 \p{t_{\ell\mu_{\t{max}}} - t_{\ell\mu_{\t{min}}}}
  \le \f12 \p{t_{\ell\mu_{\t{max}}} + \abs{t_{\ell\mu_{\t{min}}}}}
  \stackrel{\mu_{\t{max}}=0}{\le}
  \f12 t_{\ell\mu_{\t{max}}}
  + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
  \equiv \Gamma\p{t_{\ell\mu_{\t{max}}}}.
\end{align}
If is straightforward to show that $\Gamma\p{M}$ has a single maximum at $M=\sqrt{2/3}$, and that $\Gamma^\star\equiv\Gamma\p{\sqrt{2/3}}=\sqrt{3/8}$, so
\begin{align}
  g_\ell\p{p} \le \f{2\ell+1}{d}\, \gamma_\ell^2
  \stackrel{\mu_{\t{max}}=0}{\le} \p{\Gamma^\star}^2
  = \f38 < \f12.
\end{align}
Altogether, we thus find that in all cases
\begin{align}
  \bbk{\epsilon_{i\ell}^2}
  \le \f1n \times \min\Set{\f{2\ell+1}{d}, \f12}.
\end{align}
